\section*{Acknowledgments}

My friends, family, colleagues, and advisors should not be held responsible for this document's weaknesses, but they are the source of its strengths.  I thank Brian Kane and Ian Quinn for their endless patience and thoughtful engagement with strange questions.  Without their wide-ranging skepticism, guidance, and friendship, this project would not exist.  Without Dmitri Tymoczko, Dan Trueman, and Scott Burnham, I wouldn't be writing about music theory; without Jonathan Guez, Nicholas Forster, Will Sullivan, John Muniz, Sarah Weiss, Lyman Page, Angharad Davis, Becky Perry, Daniel Goldberg, Brian Miller, Jonathan Atkins, Harriett Gillham, and Edie Parrott, I wouldn't be writing; without Mary Horn Jones, my brothers, my parents, and their parents, I wouldn't be here.

The Department of Music at Yale reconfigured my engagement with the humanities, and I owe gratitude to Richard Cohn for his tireless enthusiasm and collaboration, James Hepokoski, Gundula Kreuzer, Michael Veal, Craig Wright, Daniel Harrison, Anna Zayaruznaya, Patrick McCreless, \`{E}ve Poudrier, Konrad Kaczmarek, Seth Brodsky, Sue Penney, Kristine Kinsella, and Bethany Hayes.  I owe thanks to the Allen Forte Fund, Donya Quick, Paul Hudak, Bob Frank, and especially to Alex Dubovoy, Julian Reid, and Joe McWilliams, who shared their musical work with me.

Some speech acts stick in your mind and influence the way you go about your business, and I am grateful for three in particular: to Brian Kane, who told me that a dissertation should be nothing more or less than ``an archive of your thinking;" to Salinger, who said to ``ask yourself, as a reader, what piece of writing in all the world [Andrew Jones] would most want to read if he had his heart's choice. The next step is terrible, but so simple I can hardly believe it as I write it. You just sit down shamelessly and write the thing yourself;" and to Mary, my most important collaborator, who said ``Yes."